Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints\thanks{This research was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2020-05-04 2019/NUMTA_Barkalov/SoftComputing/article.tex
New Fig1 (without decsription)
all comment fixed, but not checked

\smartqed  % flush right qed marks, e.g. at end of proof
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{hyperref,amsmath}
\usepackage[pagewise]{lineno}
\usepackage{multirow}
\usepackage[round]{natbib} 
\usepackage[misc]{ifsym}
%\usepackage{xcolor}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

@@

	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
\small
\begin{align}\label{epsilon_reserved_solution}
	\varphi(y_{\epsilon}) = \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
\end{align}
\normalsize
where the algorithm parameters $\epsilon_j, \; 1\leq j \leq m,$ are non negative numbers, and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
\small
\begin{align}\label{Y_epsilon}
	Y_{\epsilon} = \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
\end{align}
\normalsize
Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.

\begin{figure}
	\center{\includegraphics[width=0.6\linewidth]{Fig1}}
	\caption{$\varepsilon$-reserved solution.}
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y)$, $1 \leq j \leq m+1$, may be defined only partly. It means that they are defined and computable only in the sets
\small
\begin{align}\label{D_sets}
	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \; 1 \leq j \leq m.
\end{align}
\normalsize
This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined. In this approach, the first violation of any constraint indicating that the point $y$ being considered is no longer feasible terminates further computations at the point $y$.

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions in \cite{Sergeyev2006} or simplicial partitions in \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\small
\begin{align}\label{one_dimensional_problem}
	& \varphi(y(x^\ast)) =  \nonumber \\
	& \min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{align}
\normalsize
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-m}$, where the number $m>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $m=4,5$, correspondingly.
\begin{figure}
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{Fig2} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{Fig3} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $m=4$ (a), $m=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution was $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ is from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section \ref{sec:4}.

@@

	In \cite{Gergel2017_2,Gergel2019}, an approach was proposed that allows constructing the constrained global optimization problems using certain classes of multiextremal functions (in particular, the GKLS functions), in which:
\begin{itemize}
  \item the number of constraints can be varied;
  \item the volume of the feasible subdomain $Q_{m+1}$ relative to the whole search domain $D$ is defined;
	\item the global minimizer is known in advance and can be located at the boundary of the feasible domain;
	\item the number of constraints active in the global minimizer can be predefined.
\end{itemize}

@@

	The reduction of the number of trials was also confirmed by solving several series of  problems composed of the functions of the “Simple” and “Hard” classes. Each series consisted of 100 problems and was solved by the IA and IA-DL methods. The global minimizer $y^{\ast}$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e. $\| y^{\ast} - y^k \| < \varepsilon \|b - a\|$, where $a$ and $b$ are the boundaries of the search domain $D$. To reduce the dimensionality, the evolvent constructed with the parameter $m=10$ was used in both algorithms. The relative accuracy of the search for the solution at $N = 2, 3$ was $\epsilon=0.01$ and for the solution at $N=4,5$  it was $\epsilon=0.03$. The reservation parameters $\epsilon_{\nu}=\mu_{\nu} \delta,\; 1 \leq \nu \leq 2$, were set according to formula (\ref{epsilon_nu}) with $\delta=0.01$. The maximum number of iterations at $N \leq 4$ was set as $K_{max} = 10^6$, and at $N=5$ as $K_{max}=7 \cdot 10^6$.

>>>>

\smartqed  % flush right qed marks, e.g. at end of proof
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{hyperref,amsmath}
\usepackage[pagewise]{lineno}
\usepackage{multirow}
\usepackage[round]{natbib} 
\usepackage[misc]{ifsym}
\usepackage{xcolor}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

@@

	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
\small
\begin{align}\label{epsilon_reserved_solution}
	\varphi(y_{\epsilon}) = \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
\end{align}
\normalsize
where the algorithm parameters $\epsilon_j, \; 1\leq j \leq m,$ are non negative numbers, and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
\small
\begin{align}\label{Y_epsilon}
	Y_{\epsilon} = \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
\end{align}
\normalsize
Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.
{\color{red} The figure shows global minimizer $y^*$, $\epsilon$-reserved solution $y_\epsilon$ and the set $Y_\epsilon$ of all admissible points which are not worse than the $\epsilon$-reserved solution. }

\begin{figure}
	\center{\includegraphics[width=0.6\linewidth]{Fig1.pdf}}
	\caption{ {\color{red} An $\varepsilon$-reserved solution to a two-dimensional problem. } }
	\label{fig:eps_reserved_solution}
\end{figure}

@@

	We will consider the initial problem (\ref{problem}) assuming that the problem functions $g_{j}(y)$, $1 \leq j \leq m+1$, may be defined only partly. It means that they are defined and computable only in the sets
\small
\begin{align}\label{D_sets}
	D_1 = D, \; D_{j+1} = \left\{ y \in D_j: \; g_j(y) \leq 0 \right\}, \; 1 \leq j \leq m.
\end{align}
\normalsize
This situation is typical for the optimal design problems when some conditions for the functioning of the modeled object are not met and its other characteristics appear to be undefined. 
{\color{red} In this approach, the violation of any constraint at the point $y$ terminates the computation of the values of other functions at this point. }

	Many well known approaches to solving the multidimensional global optimization problems are based on either explicit or implicit reduction of multidimensional problems to one-dimensional ones (see, for example, the description of methods using the diagonal partitions in \cite{Sergeyev2006} or simplicial partitions in \cite{Zilinskas2008}). In this paper, we will use the approach based on the dimensionality reduction with the use of the space-filling curve $y(x)$ (the Peano curve). Continuous unambiguous mapping $y(x)$ projecting the interval [0,1] onto an $N$-dimensional hypercube $D$ allows reducing a multidimensional constrained minimization problem in the domain $D$ to a one-dimensional constrained optimization problem in the interval [0,1]
\small
\begin{align}\label{one_dimensional_problem}
	& \varphi(y(x^\ast)) =  \nonumber \\
	& \min \left\{\varphi(y(x)): x \in [0,1], \; g_j(y(x))\leq 0, \; 1 \leq j \leq m\right\}.
\end{align}
\normalsize
The issues of the numerical construction of the Peano space filling curve-type mappings and the corresponding theory were described in detail in \cite{Strongin2000,Sergeyev2013}. Note that the theoretical Peano curve $y(x)$ is a limit object and in practice is replaced by its approximation (called \textit{evolvent}) with predefined precision $2^{-l}$, where the integer number $l>1$ is a parameter (the \textit{density} of the evolvent). The evolvent density is determined by the accuracy of the possible practical implementation of the problem solution found. For  illustration, Fig. \ref{fig:evolvent} (a), (b) shows the projections of the interval [0,1] for the mapping $y(x)$ with $N=2$ and $l=4,5$, correspondingly.
\begin{figure}
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{Fig2} \\ {\footnotesize(a)}}
	\end{minipage}
	\hfill
	\begin{minipage}[h]{0.45\linewidth}
		\center{\includegraphics[width=1.0\linewidth]{Fig3} \\ {\footnotesize(b)}}
	\end{minipage}
	\caption{Evolvent for $N=2$ and $l=4$ (a), $l=5$ (b)}
	\label{fig:evolvent}
\end{figure}

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $l=10$ was used, the accuracy of the search for the solution was $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ is from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section \ref{sec:4}.

@@

	In \cite{Gergel2017_2,Gergel2019}, an approach was proposed that allows constructing the constrained global optimization problems using certain classes of multiextremal functions (in particular, the GKLS functions), in which:
\begin{itemize}
  \item the number of constraints can be varied;
  \item the volume of the feasible subdomain $Q_{m+1}$ relative to the whole search domain $D$ is defined;
	\item the global minimizer is known in advance and can be located at the boundary of the feasible domain;
	\item {\color{red} the number of constraints active at the global minimizer can be predefined (an inequality constraint is called active at some point if at this point the inequality turns into equality)}.
\end{itemize}

@@

	The reduction of the number of trials was also confirmed by solving several series of  problems composed of the functions of the “Simple” and “Hard” classes. Each series consisted of 100 problems and was solved by the IA and IA-DL methods. The global minimizer $y^{\ast}$ was considered to be found, if the algorithm generated a trial point $y^k$ in the vicinity of the global minimum, i.e. $\| y^{\ast} - y^k \| < \varepsilon \|b - a\|$, where $a$ and $b$ are the boundaries of the search domain $D$. To reduce the dimensionality, the evolvent constructed with the parameter $l=10$ was used in both algorithms. The relative accuracy of the search for the solution at $N = 2, 3$ was $\epsilon=0.01$ and for the solution at $N=4,5$  it was $\epsilon=0.03$. The reservation parameters $\epsilon_{\nu}=\mu_{\nu} \delta,\; 1 \leq \nu \leq 2$, were set according to formula (\ref{epsilon_nu}) with $\delta=0.01$. The maximum number of iterations at $N \leq 4$ was set as $K_{max} = 10^6$, and at $N=5$ as $K_{max}=7 \cdot 10^6$.
